Choose a License

### 2024-12-10 choose_license.qmd
Add references to licensing chapter

Some journals offer to publish your article under a Creative Commons license, but still demand an exclusive publishing and distribution license or a copyright assignment from you. This would give them more rights than the readers of the article have through the respective Creative Commons license [@Rumsey2022] and exceeds by far what is necessary to make publication possible [@Suber2022]. Consequently, authors should oppose this practice and grant publishers the same rights that every other reader of the article has. If your chosen publisher insists on an exclusive license, you may at least retain the copyright for your figures -- follow the guide "Retaining copyright for figures in academic publications to allow easy citation and reuse" by @Elson2016 to learn how to do that.

If you have published a closed-access paper before, you can consult [ShareYourPaper](https://shareyourpaper.org/) for legal options to still make it available free of charge to readers.
:::

@@

You may have noticed that we mostly refer to licenses using a name _and_ a version number. This is because the organizations that created the licenses sometimes publish updated versions to accommodate for developments in copyright law and the communities that use the licenses. For example, the Creative Commons licenses (that start with `CC`) were first published in 2002. Since then, the possibility to relicense under later ([v2.0](https://creativecommons.org/2004/05/25/announcingandexplainingournew20licenses/#:~:text=Share%20Alike%20Across%20Borders)) and compatible licenses ([v3.0](https://creativecommons.org/2007/02/23/version-30-launched/)) has been added, a 30-day window to correct license violations has been established to combat [copyleft trolls](https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling), and _sui generis_ database rights are covered explicitly ([v4.0](https://creativecommons.org/version4/)). There are many more [subtle differences between license versions](https://wiki.creativecommons.org/wiki/License_Versions), therefore it is important to indicate which license version exactly one is referring to, as the license of a work does not "update" automatically.

>>>>

Some journals offer to publish your article under a Creative Commons license, but still demand an exclusive publishing and distribution license or a copyright assignment from you. This would give them more rights than the readers of the article have through the respective Creative Commons license [@Rumsey2022] and exceeds by far what is necessary to make publication possible [@Suber2022]. Consequently, authors should oppose this practice and grant publishers the same rights that every other reader of the article has. To facilitate self-archiving, one can also modify the contract with publishers via a rights retention statement [@UKRN2023] / author's addendum [@SPARC2006]. If you have published a closed-access paper before, you can consult [ShareYourPaper](https://shareyourpaper.org/) for legal options to still make it available free of charge to readers.

If your chosen publisher insists on an exclusive license, you may at least retain the copyright for your figures -- follow the guide "Retaining copyright for figures in academic publications to allow easy citation and reuse" by @Elson2016 to learn how to do that.
:::

@@

You may have noticed that we mostly refer to licenses using a name _and_ a version number. This is because the organizations that created the licenses sometimes publish updated versions to accommodate for developments in copyright law and the communities that use the licenses. For example, the Creative Commons licenses (that start with `CC`) were first published in 2002. Since then, the possibility to relicense under later ([v2.0](https://creativecommons.org/weblog/entry/4216#:~:text=Share%20Alike%20Across%20Borders), 2004) and compatible licenses ([v3.0](https://creativecommons.org/weblog/entry/7249), 2007) has been added, a 30-day window to correct license violations has been established to combat [copyleft trolls](https://commons.wikimedia.org/wiki/Commons:Copyleft_trolling), and _sui generis_ database rights are covered explicitly ([v4.0](https://creativecommons.org/weblog/entry/40768), 2013). There are many more [subtle differences between license versions](https://wiki.creativecommons.org/wiki/License_Versions), therefore it is important to indicate which license version exactly one is referring to, as the license of a work does not "update" automatically. CC0, officially introduced in 2009, is still at version [1.0](http://www.creativecommons.org/weblog/entry/13304).
